---
title: "SQL"
output: html_document
---

## Filtering Columns and Rows


with SQL, you can filter columns with SELECT and rows with WHERE. Let's try an example with Lahman baseball database. First we need to load the appropriate packages. 


```{r warning=FALSE, message=FALSE}
library(Lahman)
library (sqldf)
```

#______________________________________________________________________________

####Example 1

Now, suppose we would like to see homerun totals for the 1927 Yankees. We could write the following:


```{r}
query<-"SELECT playerID, yearID, teamID, HR FROM Batting
WHERE teamID='NYA' and yearID=1927"
sqldf(query)
```

#___________________________________________________________________________________

####Example 2

Now, we can list all players from Yankees that have scored at least 40 homeruns:


```{r}
query<-"SELECT playerID,teamID,yearID,HR FROM Batting WHERE teamID='NYA' and HR>=40"
sqldf(query)
```

#___________________________________________________________________________________

####Example 3

We can also see those players who had 40 homeruns but didn't have more than 60 strikeouts:


```{r}
query<-"SELECT playerID,teamID,yearID,HR,SO FROM Batting WHERE HR>40 and SO<60"
sqldf(query)
```

#___________________________________________________________________________________

####Example 4

Here is the information from 1970's for Phillies for all players who had more than 30 homeruns:


```{r}
query<-"SELECT playerID, teamID,yearID,HR FROM Batting WHERE yearID>=1970 and yearID<1980 and HR>30 and teamID='PHI'"
sqldf(query)
```

#___________________________________________________________________________________


###Order By
####Example 1

We can also arrange them by descending number homeruns: 


```{r}
query<-"SELECT playerID,teamID,yearID,HR FROM Batting WHERE HR>50
ORDER BY HR DESC"
sqldf(query)
```

#___________________________________________________________________________________

####Example 2

This is the list of players that had at least 400 at bats but had less than 10 strikeouts and it is arranged by the least amount of strikeouts


```{r}
query<-"SELECT playerID,teamID,yearID,SO,AB FROM Batting 
WHERE SO<10 and AB>=400
ORDER BY SO"
sqldf(query)
```

#___________________________________________________________________________________

###Aggregation
####Example 1
Here we can see the players who had more than 600 homeruns:


```{r}
query<-"SELECT playerID,sum(HR) FROM Batting
GROUP BY playerID
HAVING sum(HR)>=600
 ORDER BY sum(HR) DESC"
sqldf(query)
```

#___________________________________________________________________________________

####Example 2
Here is the list of players that have avarage of more than 30 homeruns and order them by descending order:


```{r}
query<-"SELECT playerID,avg(HR) FROM Batting
GROUP BY playerID
HAVING avg(HR)>30
ORDER BY avg(HR) DESC"
sqldf(query)
```

#___________________________________________________________________________________

###Joins
###Example 1
Here is an example of joining two tables:


```{r}
query<-"SELECT nameFirst,nameLast,teamID,yearID,HR 
FROM Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
WHERE Batting.playerID='ruthba01'"
sqldf(query)
```

#___________________________________________________________________________________

###Example 2
Here is an example how to order joined tables results from most to least homeruns:


```{r}
query<-"SELECT nameFirst,nameLast,teamID,yearID,HR 
FROM Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
WHERE HR>50
ORDER BY HR DESC"

sqldf(query)
```

#___________________________________________________________________________________

###Example 3
Here, we can see how to join with table "teams":


```{r}
query<-"SELECT playerID,name,Batting.yearID,Batting.HR 
From Batting INNER JOIN Teams
ON Batting.teamID=Teams.teamID AND Batting.yearID=Teams.yearID
WHERE playerID='ruthba01'"


sqldf(query)
```

#___________________________________________________________________________________

###Example 4
Here is a list of all players who had more than 1 milion dollars average salary and have a first name Bob:

```{r}
query<-"SELECT Master.playerID,nameFirst,nameLast,yearID,avg(salary)
FROM Master INNER JOIN Salaries
on Master.playerID=salaries.playerID
GROUP BY Master.playerID
Having avg(salary)>=1000000 and nameFirst='Bob'
ORDER BY avg(salary) DESC"

sqldf(query)
```

#___________________________________________________________________________________